---
title: "gbifniche"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


R tool for assessing uncertainties around qualities of observation for environmental niches using GBIF data


## Requirements

+ [R v4.1](https://www.r-project.org/)
+ [rlang](https://CRAN.R-project.org/package=rlang)
+ [dplyr](https://CRAN.R-project.org/package=dplyr)
+ [readr](https://CRAN.R-project.org/package=readr)
+ [ggplot2](https://CRAN.R-project.org/package=ggplot2)
+ [rgbif](https://CRAN.R-project.org/package=rgbif)
+ [rappdirs](https://CRAN.R-project.org/package=rappdirs)
+ [mapr](https://CRAN.R-project.org/package=mapr)

## Installation

```
remotes::install_github("kolive4/gbif_thermal_niche")
```

## Usage

Retrieve a data set by species.

```{r}
suppressPackageStartupMessages({
  library(gbifniche)
  library(dplyr)
})

# make it quiet for the purpose of the README markdown
x = fetch_gbif(species = "Squalus acanthias subsp. acanthias Linnaeus, 1758", 
               verbose = FALSE, progress = FALSE) |>
  dplyr::glimpse()
```

If you have downloaded multiple datasets you can list the ones you have.

```{r}
(spp <- list_gbif())
```

Now that you have a set of data files you can read them, optionally with a select number of [GBIF Darwin Core](https://ipt.gbif.org/manual/en/ipt/2.5/occurrence-data) required and recommended fields.

```{r}
x <- read_gbif(spp[3], dwc = TRUE)
glimpse(x)
```

And plot...

```{r}
plot_gbif(x, what = 'ggplot')
```